Data Exploration

### 2019-03-14 src/data_exploration.Rmd
playing with maxent variable selection

```{r}
library(dismo)
library(rJava)
jar <- paste(system.file(package="dismo"), "C:/Users/bunny/Documents/GitHub/Chapter_1/packrat/lib/x86_64-w64-mingw32/3.5.1/dismo/java/maxent.jar", sep='')
file.exists(jar)
testmax <- maxent(janpresabmax, janpresabmax$occurrence)

>>>>

# maxentvariableselection package

https://cran.r-project.org/web/packages/MaxentVariableSelection/MaxentVariableSelection.pdf

### 2019-03-14 src/stack_test.Rmd
playing with maxent variable selection

pre5 <- predict(mod5, test, args='outputformat=logistic' ) 
preplot5 <- plot(pre5)
preplot5
```

>>>>

pre5 <- predict(mod5, test, args='outputformat=logistic' ) 
preplot5 <- plot(pre5)
preplot5

#-----------------------

mod6 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "replicates=5", "replicatetype=crossvalidate")) #replicates with crossvalidate result in multiple map outputs (one for each replicate). It is not possible to 'merge' these maps https://stackoverflow.com/questions/28885987/r-how-to-get-one-summary-prediction-map-instead-for-5-when-using-5-fold-cros
mod6

pre6 <- predict(mod6, test, args='outputformat=logistic' ) 
preplot6 <- plot(pre6)
preplot6

#-----------------------
strt <- Sys.time() #get the start time
mod7 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) #threads should speed things up...
mod7

pre7 <- predict(mod7, test, args='outputformat=logistic' ) 
preplot7 <- plot(pre7)
preplot7
print(Sys.time()-strt) #time it took to run

#43 secs

strt <- Sys.time() #get the start time
mod7.2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE"))
mod7.2

pre7.2 <- predict(mod7.2, test, args='outputformat=logistic' ) 
preplot7.2 <- plot(pre7)
preplot7.2
print(Sys.time()-strt) #time it took to run

#1.75 mins

#conf that threads does indeed speed up process

#---------------------
#follwoing radosavljevic 2014

mod8.1 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=0.25")) #playing with multiplier
mod8.1

pre8.1 <- predict(mod8.1, test, args='outputformat=logistic' ) 
preplot8.1 <- plot(pre8.1)
preplot8.1

mod8 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) #default setting for betam is 1
mod8

pre8 <- predict(mod8, test, args='outputformat=logistic' ) 
preplot8 <- plot(pre8)
preplot8

mod8.4 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=1.25")) #playing with multiplier
mod8.4

pre8.4 <- predict(mod8.4, test, args='outputformat=logistic' ) 
preplot8.4 <- plot(pre8.4)
preplot8.4

mod8.2 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=2")) 
mod8.2

pre8.2 <- predict(mod8.2, test, args='outputformat=logistic' ) 
preplot8.2 <- plot(pre8.2)
preplot8.2

mod8.3 <- maxent(x = prab, p = pa, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4", "betamultiplier=6")) 
mod8.3

pre8.3 <- predict(mod8.3, test, args='outputformat=logistic' ) 
preplot8.3 <- plot(pre8.3)
preplot8.3

one maxent run with all variables

```{r}
presaball <- read.csv("../output/bio/presab_10k.csv", header = TRUE)
janpresaball <- subset(presaball, month == 1)
janpresaball <- subset(janpresaball, select = c(occurrence, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
paall <- janpresaball$occurrence
janpresaball <- subset(janpresaball, select = -c(occurrence))

temp_depth <- raster(stacklist$temp_depth)
crs(temp_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(temp_depth) <- "temp_depth"
temp_depth <- projectRaster(temp_depth, aea)
temp_surface <- raster(stacklist$temp_surface)
crs(temp_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(temp_surface) <- "temp_surface"
temp_surface <- projectRaster(temp_surface, aea)
chl_depth <- raster(stacklist$chl_depth)
crs(chl_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(chl_depth) <- "chl_depth"
chl_depth <- projectRaster(chl_depth, aea)
chl_surface <- raster(stacklist$chl_surface)
crs(chl_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(chl_surface) <- "chl_surface"
chl_surface <- projectRaster(chl_surface, aea)
o2_depth <- raster(stacklist$o2_depth)
crs(o2_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(o2_depth) <- "o2_depth"
o2_depth <- projectRaster(o2_depth, aea)
o2_surface <- raster(stacklist$o2_surface)
crs(o2_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(o2_surface) <- "o2_surface"
o2_surface <- projectRaster(o2_surface, aea)
salinity_depth <- raster(stacklist$salinity_depth)
crs(salinity_depth) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(salinity_depth) <- "salinity_depth"
salinity_depth <- projectRaster(salinity_depth, aea)
salinity_surface <- raster(stacklist$salinity_surface)
crs(salinity_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(salinity_surface) <- "salinity_surface"
salinity_surface <- projectRaster(salinity_surface, aea)
ssh_surface <- raster(stacklist$ssh_surface)
crs(ssh_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(ssh_surface) <- "ssh_surface"
ssh_surface <- projectRaster(ssh_surface, aea)
mlp_surface <- raster(stacklist$mlp_surface)
crs(mlp_surface) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(mlp_surface) <- "mlp_surface"
mlp_surface <- projectRaster(mlp_surface, aea)
nao_sample <- raster(stacklist$nao_sample)
crs(nao_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(nao_sample) <- "nao_sample"
nao_sample <- projectRaster(nao_sample, aea)
nao_prev <- raster(stacklist$nao_prev)
crs(nao_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(nao_prev) <- "nao_prev"
nao_prev <- projectRaster(nao_prev, aea)
nao_winter <- raster(stacklist$nao_winter)
crs(nao_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
names(nao_winter) <- "nao_winter"
nao_winter <- projectRaster(nao_winter, aea)
amo_sample <- raster(stacklist$amo_sample)
crs(amo_sample) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(amo_sample) <- "amo_sample"
amo_sample <- projectRaster(amo_sample, aea)
amo_prev <- raster(stacklist$amo_prev)
crs(amo_prev) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0" 
names(amo_prev) <- "amo_prev"
amo_prev <- projectRaster(amo_prev, aea)
amo_winter <- raster(stacklist$amo_winter)
crs(amo_winter) <- "+proj=longlat +datum=WGS84 +ellps=WGS84 +towgs84=0,0,0"
names(amo_winter) <- "amo_winter"
amo_winter <- projectRaster(amo_winter, aea)

testall <- stack(temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter)

mod1all <- maxent(x = janpresaball, p = paall, args=c("maximumiterations=5000", "responsecurves=TRUE", "jackknife=TRUE", "threads=4")) 
mod1all

pre1all <- predict(mod1all, testall, args='outputformat=logistic' ) 
preplot1all <- plot(pre1all)
preplot1all

# maxentvariableselection package

https://cran.r-project.org/web/packages/MaxentVariableSelection/MaxentVariableSelection.pdf + https://cran.r-project.org/web/packages/MaxentVariableSelection/vignettes/MaxentVariableSelection.pdf

```{r}
janaball <- read.csv("../output/bio/background_10k.csv", header = TRUE)
janpresall <- read.csv("../output/bio/pres_nomissing.csv", header = TRUE)
janpresaball <- rbind.fill(janaball, janpresall)
janpresaball <- subset(janpresaball, month == 1)
janpresaball <- subset(janpresaball, select = c(occurrence, longitude_meters, latitude_meters, temp_depth, temp_surface, chl_depth, chl_surface, o2_depth, o2_surface, salinity_depth, salinity_surface, ssh_surface, mlp_surface, nao_sample, nao_prev, nao_winter, amo_sample, amo_prev, amo_winter))
names(janpresaball)[names(janpresaball)=="occurrence"] <- "species"
names(janpresaball)[names(janpresaball)=="longitude_meters"] <- "longitude"
names(janpresaball)[names(janpresaball)=="latitude_meters"] <- "latitude"
janpresall <- subset(janpresaball, species == 1)
janpresall$species <- replace(janpresall$species, janpresall$species == "1", "fish")
write.csv(janpresall, "../output/maxentvariableselection/presence.csv", row.names = FALSE)
janaball <- subset(janpresaball, species == 0)
janpresall$species <- replace(janpresall$species, janpresall$species == "0", "bg")
write.csv(janaball, "../output/maxentvariableselection/background.csv", row.names = FALSE)

VariableSelection(maxent="C:/Users/bunny/Downloads/maxent/maxent.jar", outdir=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/output"), gridfolder=("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/testgrid"), occurrencelocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/presence.csv"), backgroundlocations= ("C:/Users/bunny/Documents/GitHub/Chapter_1/output/maxentvariableselection/background.csv"), additionalargs="nolinear noquadratic noproduct nothreshold noautofeature", contributionthreshold=5, correlationthreshold=0.9, betamultiplier=seq(0.5, 4, 0.25))
```
